Kids PM Summary

### 2019-12-09 data_cleaning_initial.Rmd
dusttrak cleaning and summary stats

---
title: "KidsAIR initial data work"
author: "Ethan Walker"
date: "Started 19 Nov 2019, Updated 5 Dec 2019"
output: pdf_document
---

@@

# Save pm dataset as RDS and CSV
write_rds(kids_pm, "Output/kids_pm_full.rds")
#write_csv(kids_pm, "Output/kids_pm_full.csv")

# Load data and clean up
kids_pm <- read_rds("Output/kids_pm_full.rds")

kids_pm_check <- kids_pm %>% 
  group_by(home_winter_id) %>% 
  summarise(obs = n()) %>% 
  ungroup()

ak_dt_home <- read_csv("Input/AK/ak_dt_home.csv")
nn_dt_home <- read_csv("Input/NN/nn_dt_home.csv")
wmt_dt_home <- read_csv("Input/WMT/wmt_dt_home.csv")

dt_home <- rbind(ak_dt_home, nn_dt_home, wmt_dt_home) %>% 
  rename_all(tolower) %>% 
  filter(!is.na(homewinterid)) %>% 
  mutate(home_winter_id = as.character(homewinterid)) %>% 
  left_join(kids_pm_check, by = "home_winter_id")

##### Add to new variables to Kids PM data #####
kids_pm_new <- kids_pm %>% 
  mutate(home_winter_id = as.character(home_winter_id),
         area = as.factor(area),
         pm_date = date,
         pm_time = time,
         pm_datetime = datetime,
         pm = as.numeric(pm)) %>% 
  select(home_winter_id, area, pm_date, pm_time, pm_datetime, pm) %>% 
  arrange(home_winter_id, pm_date) %>% 
  group_by(home_winter_id) %>% 
  # new variables summarizing pm2.5 within each home_winter_id
  mutate(pm_total_observations = n(),
         pm_total_summed = sum(pm),
         min_datetime = first(pm_datetime),
         max_datetime = last(pm_datetime),
         min_datetime = ymd_hms(min_datetime),
         max_datetime = ymd_hms(max_datetime),
         diff_datetime = interval(min_datetime, max_datetime),
         pm_total_days = diff_datetime/86400,
         pm_24hr_mean = pm_total_summed/pm_total_days,
         pm_rolling_24hr = rollsum(pm, 1440, fill = "extend", align = "right"),
         obs = row_number(),
         sampling_day = if_else(obs < 1441, 1, 0),
         sampling_day = if_else(obs < 2881 & obs > 1440, 2, sampling_day),
         sampling_day = if_else(obs < 4321 & obs > 2880, 3, sampling_day),
         sampling_day = if_else(obs < 5761 & obs > 4320, 4, sampling_day),
         sampling_day = if_else(obs < 7201 & obs > 5760, 5, sampling_day),
         sampling_day = if_else(obs < 8641 & obs > 7200, 6, sampling_day),
         sampling_day = if_else(obs < 10081 & obs > 8640, 7, sampling_day),
         sampling_day = if_else(obs > 10080, 8, sampling_day)) %>%
  group_by(home_winter_id, sampling_day) %>% 
  mutate(pm_24hr_daily = sum(pm),
         pm_24hr_daily_weighted = pm_24hr_daily*(1440/n()),
         pm_sampling_day = sampling_day) %>% 
  ungroup() %>% 
  select(-min_datetime, -max_datetime, -diff_datetime, -obs, -sampling_day)

>>>>

---
title: "KidsAIR initial data work"
author: "Ethan Walker"
date: "Started 19 Nov 2019, Updated 9 Dec 2019"
output: pdf_document
---

@@

# Save pm dataset as RDS and CSV
write_rds(kids_pm, "Output/kids_pm_full_raw.rds")
#write_csv(kids_pm, "Output/kids_pm_full.csv")
```

# Load Dusttrak data and clean up
```{r}
# Load full, raw dataset compiled in chunk above
kids_pm_full <- read_rds("Output/kids_pm_full_raw.rds")

# Load and combine data sheets with QAQC comments, etc
ak_dt_home <- read_csv("Input/AK/ak_dt_home.csv") %>% 
  mutate(area = "AK")
ak_dt <- read_xlsx("Input/AK/ak_dustrak.xlsx") %>% 
  mutate(area = "AK") %>% 
  rename_all(tolower) %>% 
  select(homewinterid, dtid, area)
nn_dt_home <- read_csv("Input/NN/nn_dt_home.csv") %>% 
  mutate(area = "NN")
nn_dt <- read_xlsx("Input/NN/nn_dustrak.xlsx") %>% 
  mutate(area = "NN") %>% 
  rename_all(tolower) %>% 
  select(homewinterid, dtid, area)
wmt_dt_home <- read_csv("Input/WMT/wmt_dt_home.csv") %>% 
  mutate(area = "WMT")
wmt_dt <- read_xlsx("Input/WMT/wmt_dustrak.xlsx") %>% 
  mutate(area = "WMT") %>% 
  rename_all(tolower) %>% 
  select(homewinterid, dtid, area)

dt <- rbind(ak_dt, nn_dt, wmt_dt) %>% 
  mutate(home_winter_id = as.character(homewinterid)) %>% 
  select(-homewinterid)

dt_home <- rbind(ak_dt_home, nn_dt_home, wmt_dt_home) %>% 
  rename_all(tolower) %>% 
  filter(!is.na(homewinterid)) %>% 
  mutate(home_winter_id = as.character(homewinterid)) %>% 
  select(-homewinterid) %>% 
  left_join(dt, by = c("area", "home_winter_id"))

# Join dt_home with full dataset
kids_pm_joined <- kids_pm_full %>% 
  left_join(dt_home, by = c("area", "home_winter_id")) %>% 
  rename(treatment = condition)

##### Add to new variables to Kids PM data #####
kids_pm_new <- kids_pm_joined %>%  
  #filter(area == "AK") %>% 
  # apply correction factor to pm and change to ug
  mutate(pm = pm/1.65,
         pm = pm*1000) %>% 
  mutate(home_winter_id = as.character(home_winter_id),
         area = as.factor(area),
         pm_date = date,
         pm_time = time,
         pm = as.numeric(pm),
         day_of_week = weekdays(date)) %>% 
  unite("pm_datetime", date:time, sep = " ") %>% 
  #select(home_winter_id, area, pm_date, pm_time, pm_datetime, day_of_week, pm) %>% 
  arrange(area, home_winter_id, pm_date) %>% 
  group_by(area, home_winter_id) %>% 
  # new variables summarizing pm2.5 within each home_winter_id
  mutate(pm_total_observations = n(),
         # new datetime values to calculate sampling interval
         min_datetime = first(pm_datetime),
         max_datetime = last(pm_datetime),
         min_datetime = ymd_hms(min_datetime),
         max_datetime = ymd_hms(max_datetime),
         diff_datetime = interval(min_datetime, max_datetime),
         # total sampling interval, based on first and last datetimes
         pm_sample_interval = diff_datetime/86400,
         # expected number of observations with the sampling interval
         expected_obs = pm_sample_interval*24*60,
         # difference between actual and expected observations
         actual_vs_expected_diff = pm_total_observations - expected_obs,
         # indicator variable if actual vs expected obs > 10 or greater
         actual_vs_expected = if_else(actual_vs_expected_diff < 10 &
                                      actual_vs_expected_diff > -10, 1, 0),
         pm_mean = mean(pm),
         # breaking sampling period into days
         obs = row_number(),
         sampling_day = if_else(obs < 1441, 1, 0),
         sampling_day = if_else(obs < 2881 & obs > 1440, 2, sampling_day),
         sampling_day = if_else(obs < 4321 & obs > 2880, 3, sampling_day),
         sampling_day = if_else(obs < 5761 & obs > 4320, 4, sampling_day),
         sampling_day = if_else(obs < 7201 & obs > 5760, 5, sampling_day),
         sampling_day = if_else(obs < 8641 & obs > 7200, 6, sampling_day),
         sampling_day = if_else(obs < 10081 & obs > 8640, 7, sampling_day),
         sampling_day = if_else(obs > 10080, 8, sampling_day)) %>%
  # calculate pm means for each sampling day
  group_by(area, home_winter_id, sampling_day) %>% 
  mutate(pm_mean_daily = mean(pm),
         pm_sampling_day = sampling_day) %>% 
  ungroup() %>% 
  select(-min_datetime, -max_datetime, -diff_datetime, -obs, -sampling_day, -datetime)

### 2019-12-09 pm_summary.Rmd
dusttrak cleaning and summary stats

# Summary stats
```{r}
summary_kids_pm <- kids_pm %>% 
  na.exclude() %>% 
  #filter(pm < 0)
  summarize(mean(pm_24hr_mean), sd(pm_24hr_mean),
            min(pm_24hr_mean), max(pm_24hr_mean),
            min(pm), max(pm), mean(pm_total_days),
            min(pm_total_days), max(pm_total_days),
            min(pm_total_observations), max(pm_total_observations),
            mean(pm_24hr_daily), mean(pm_24hr_daily_weighted)) %>% 
  t()
summary_kids_pm
```

# Boxplots
```{r, fig.height=6, fig.width=5}
boxplots_kids_pm <- kids_pm %>% 
  ggplot(aes(area, pm_24hr_mean)) +
    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
               fatten = 1, outlier.size = 1.5, width = 0.35) +
    stat_summary(fun.y=mean, geom="point", shape=17, size=3, color="red") +
    theme_minimal() +
    ylim(0, 400) +
    #geom_hline(yintercept = 50, color = "red", size = 1.2) +
    labs(y = expression(paste("PM"[2.5], " (", mu, g/m^3, ")"))) +
    #labs(aes(x = "", y = paste("48-hour indoor PM "[2.5], " (", mu, g/m^3, ")"))) +
    theme(axis.text.x = element_text(size = 16, colour = "black"),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank())
boxplots_kids_pm
```

>>>>

# Summary stats
```{r}
# summary by area
summary_area <- kids_pm %>%
  #filter(actual_vs_expected == 1) %>% 
  #filter(pm <0) %>% 
  group_by(home_winter_id) %>% 
  summarize("n" = n(),
            "Mean PM" = mean(pm, na.rm = TRUE), 
            "SD PM" = sd(pm, na.rm = TRUE),
            "Min PM" = min(pm, na.rm = TRUE), 
            "Median PM" = median(pm, na.rm = TRUE),
            "Max PM" = max(pm, na.rm = TRUE),
            "Mean Sample Days" = mean(pm_sample_interval, na.rm = TRUE),
            "Min Sample Days" = min(pm_sample_interval, na.rm = TRUE), 
            "Max Sample Days" = max(pm_sample_interval, na.rm = TRUE),
            "Mean Sample Obs" = mean(pm_total_observations, na.rm = TRUE),
            "Min Sample Obs" = min(pm_total_observations, na.rm = TRUE), 
            "Max Sample Obs" = max(pm_total_observations, na.rm = TRUE)) %>% 
  t()
summary <- as.data.frame(summary_area)
my_names <- summary_area[1, ]
colnames(summary) <- my_names
summary_area <- summary %>% 
  rownames_to_column() %>% 
  rename(Statistic = rowname) %>% 
  filter(Statistic != "area") 
#summary_area

# summary by home_winter_id
summary_id <- kids_pm %>% 
  group_by(area, home_winter_id) %>% 
  summarise("n" = n(),
            "Mean PM" = mean(pm, na.rm = TRUE), 
            "SD PM" = sd(pm, na.rm = TRUE),
            "Min PM" = min(pm, na.rm = TRUE), 
            "Median PM" = median(pm, na.rm = TRUE),
            "Max PM" = max(pm, na.rm = TRUE),
            "Comments" = first(qaqccomments))

# summary by dusttrak id
summary_dtid <- kids_pm %>% 
  group_by(area, dtid) %>% 
  summarise("n" = n(),
            "Mean PM" = mean(pm, na.rm = TRUE), 
            "SD PM" = sd(pm, na.rm = TRUE),
            "Min PM" = min(pm, na.rm = TRUE), 
            "Median PM" = median(pm, na.rm = TRUE),
            "Max PM" = max(pm, na.rm = TRUE))
```

# Boxplots
```{r, fig.height=6, fig.width=5}
boxplots_kids_pm <- kids_pm %>% 
  filter(pm < 500) %>% 
  ggplot(aes(area, pm)) +
    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
               fatten = 1, outlier.size = 1.5, width = 0.35) +
    stat_summary(fun.y=mean, geom="point", shape=17, size=3, color="red") +
    theme_minimal() +
    ylim(-100, 500) +
    #geom_hline(yintercept = 50, color = "red", size = 1.2) +
    labs(y = expression(paste("PM"[2.5], " (", mu, g/m^3, ")"))) +
    #labs(aes(x = "", y = paste("48-hour indoor PM "[2.5], " (", mu, g/m^3, ")"))) +
    theme(axis.text.x = element_text(size = 16, colour = "black"),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank())
boxplots_kids_pm
```
